---
output: 
  word_document:
    reference_docx: "phs-offdev-summary.docx"
    fig_caption: no
---


```{r setup, include = FALSE}

library(ggplot2)

```

```{r Total referrals data, include = FALSE}
figs_referrals <- read_parquet(
  paste0("//PHI_conf/MentalHealth5/CAPTND/CAPTND_shorewise/output/analysis_", data_analysis_latest_date,   "/shorewise_publication/data/referrals/table_referrals_quarterly.parquet")) |> 
  filter(dataset_type == dataset_choice,
         `Health board` == "NHS Scotland") |> 
  select("last_yr" = 3, "last_qt" = 6, "current_qt" = 7) |> 
  mutate(across(everything(),~ gsub(",", "", .))) |> 
  mutate_all(as.numeric) |> 
  mutate(diff_qt = current_qt - last_qt,
         diff_yr = current_qt - last_yr,
         prop_qt = round(100-(last_qt/current_qt*100), 1),
         prop_yr = round(100-(last_yr/current_qt*100), 1),
         across(1:5, ~prettyNum(., big.mark = ",")))
```

```{r Referral acceptance data, include = FALSE}
table_data2 <- read_parquet(
  paste0("//PHI_conf/MentalHealth5/CAPTND/CAPTND_shorewise/output/analysis_", data_analysis_latest_date, "/shorewise_publication/data/non_acceptance/table_acc_rate.parquet")) |> 
  ungroup() |> 
  filter(dataset_type == dataset_choice) |> 
  select(-dataset_type)

```

```{r Appointments data, include = FALSE}
table_data3 <- read_parquet(
  paste0("//PHI_conf/MentalHealth5/CAPTND/CAPTND_shorewise/output/analysis_", data_analysis_latest_date, "/shorewise_publication/data/appointments_att/table_apps_att_latest_qt.parquet")) |> 
  filter(dataset_type == dataset_choice) |> 
  select(-dataset_type)

```

<Div custom-style = "Publication Title">Child, Adolescent, and Psychological Therapies National Dataset (CAPTND):</Div>
<Div custom-style = "Publication subtitle">`r paste0(dataset_label, " Report")`</Div>
<Div custom-style = "Stats designation">An Official statistics in development release for Scotland</Div>
<Div custom-style = "Publication date">Publication date: 03 September 2024</Div>
<Div custom-style = "Restricted Stats">RESTRICTED STATISTICS: embargoed to 09:30 dd/mm/yyyy</Div>

# About this release
This release by Public Health Scotland (PHS) provides an update on information relating to 
activity in `r dataset_label_long` `r ifelse(dataset_choice == "PT", "services", "")`
from the Child, Adolescent, and Psychological Therapies National Dataset (CAPTND) in
NHSScotland, for the quarter ending `r date_label`.

# Main Points
For `r dataset_choice` in the quarter ending `r date_label`:

<Div custom-style = "Bullet 1"> 
There were a total of `r figs_referrals$current_qt` `r ifelse(dataset_choice == "CAMHS", "children and young people", "people")` 
referred to `r dataset_choice` recorded in CAPTND. This is `r ifelse(figs_referrals$diff_qt > 0, "an increase", "a decrease")` of 
`r figs_referrals$diff_qt` (`r figs_referrals$prop_qt`%) compared with the previous quarter, 
and `r ifelse(figs_referrals$diff_yr > 0, "an increase", "a decrease")` of `r figs_referrals$diff_yr` 
(`r figs_referrals$prop_yr`%) compared with the same quarter ending `r date_label_yr`. </Div>

<Div custom-style = "Bullet 1">
Of the `r table_data2[[15,5]]` referrals to `r dataset_choice`, `r table_data2[[15,2]]`
(`r table_data2[[15,6]]`) were accepted. </Div>

<Div custom-style = "Bullet 1">
There were a total of `r table_data3[[15,2]]` appointments in `r dataset_choice`
across NHS Scotland. Of these, `r table_data3[[15,3]]` were first contact appointments. </Div>

<Div custom-style = "Table or chart caption">`Chart title`</Div>
```{r chart, fig.width=7.18, fig.height=3.31, fig.cap="Appropriate ALT Text to be added."}
chart
```

# Background

Waiting times information for Child and Adolescent Mental Health Services (CAMHS) and 
psychological therapies using CAPTND is still being developed. NHS Boards are working with 
PHS and the Scottish Government to improve the consistency and completeness of the 
information.  

if(dataset_choice == "CAMHS){

"CAMHS are multi-disciplinary teams that provide treatment and/or interventions for children and
young people experiencing mental health problems. Where a child or young person is awaiting
an assessment for a neurodevelopmental condition such as ASD, or ADHD, they may not be
appropriate for CAMHS. This release does not report on the neurodevelopmental (ND) pathway
including those children and young people on the neurodevelopmental waiting list."

}else{

"Psychological therapies refer to a range of interventions, based on psychological concepts and
theory, which are designed to help people understand and make changes to their thinking,
behaviour and relationships in order to relieve distress and to improve functioning. The standard
applies specifically to psychological therapies for treatment of a mental illness or disorder."
}

# Contact
Name | Role | Number | [phs.generic-team-email@phs.scot](mailto:phs.generic-team-email@phs.scot)

Name | Role | Number | [phs.generic-team-email@phs.scot](mailto:phs.generic-team-email@phs.scot)

For all media enquiries please email [phs.comms@phs.scot](mailto:phs.comms@phs.scot) or call 0131 275 6105.

# Further information
Data from this publication are available from the [publication page](https://publichealthscotland.scot/publications/YOUR-PUBLICATION-URL) on our website.

Open data from this publication is available from the [Scottish Health and Social Care Open Data platform](https://www.opendata.nhs.scot/).

Our statistical practice is regulated by the Office for Statistics Regulation (OSR). Official Statistics in development are official statistics which are undergoing a development. 

Visit our website for [further information about our statistics and PHS as an Official Statistics producer.](https://publichealthscotland.scot/our-organisation/about-our-statistics/official-statistics/)

The next release of this publication will be DD Month YYYY.

Other formats of this publication are available on request by calling 0131 314 5300 or emailing [phs.otherformats@phs.scot](mailto:phs.otherformats@phs.scot).

# NHS Performs
A selection of information from this publication is included in [NHS Performs](http://www.nhsperforms.scot). NHS Performs is a website that brings together a range of information on how hospitals and NHS Boards within NHSScotland are performing.
